---
title: "Alluvial_plot_TCRbulk_MA"
author: "Quy Ai Ngo"
date: "09-Nov-2021"
output: slides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r setup, echo = TRUE}
setwd("/home/localadmin/Desktop/RESEARCH/MArnaud/bulk_TCR")

library(data.table)
library(tidyverse)
library(ggplot2)
library(ggforce)
#library(gg.gap)
#library(ggbreak)
library(ggalluvial)
```


```{r data import & preparation, echo=FALSE}
# Get a List of all files named with a key word, say all `.14MH` files
filenames <- list.files("./14MH/TCR_repertoire",
                        pattern = "*.txt", full.names=TRUE)
filenames

# Load data sets
list.DF <- lapply(filenames, data.table::fread)
lapply(list.DF, dim)

# change 1st column name to get rid of the character "#" & subset dataframes with only IN frame clones
lapply(list.DF, setnames, "#CDR3_sequence", "CDR3_sequence")

# subset DF list to 1read, 2read & INframe list
list.DF.IN <- lapply(list.DF, subset, Frame == "IN")
list.DF.1r <- lapply(list.DF, subset, Count == 1)
list.DF.1rIN <- lapply(list.DF, subset, Frame == "IN" & Count == 1)
#list.DF.2rIN <- lapply(list.DF, subset, Frame == "IN" & Count == 2)

```


```{r count clonotypes for each list of df}
#DEV!!! the for loop does NOT work

source("/home/localadmin/Desktop/RESEARCH/MArnaud/bulk_TCR/clonotypeCount.R")

for (i in c(list.DF, list.DF.1r, list.DF.1rIN, list.DF.2rIN)) {
  print(paste("data frame = ", deparse(substitute(i))))
  count.list.DF <- lapply(i, clonotypeCount)
}

#print(paste("data frame = ", deparse(substitute(list.DF.1r))))
#lapply(list.DF.1r, clonotypeCount)

#print(paste("data frame = ", deparse(substitute(list.DF.1rIN))))
#lapply(list.DF.1rIN, clonotypeCount)

#print(paste("data frame = ", deparse(substitute(list.DF.2rIN))))
#lapply(list.DF.2rIN, clonotypeCount)

```


```{r name DF in each list, echo=FALSE}

nameDF = c("p14MH.1_B0", "p14MH.2_T0",
           "p14MH.3_preREP", "p14MH.4_ACTP",
           "p14MH.5_B14", "p14MH.6_T41")

nameDF.1r = c("p14MH.1r_B0", "p14MH.1r_T0",
              "p14MH.1r_preREP", "p14MH.1r_ACTP",
              "p14MH.1r_B14", "p14MH.1r_T41")

nameDF.IN = c("p14MH.IN_B0", "p14MH.IN_T0",
              "p14MH.IN_preREP", "p14MH.IN_ACTP",
              "p14MH.IN_B14", "p14MH.IN_T41")

nameDF.1rIN = c("p14MH.1rIN_B0", "p14MH.1rIN_T0",
              "p14MH.1rIN_preREP", "p14MH.1rIN_ACTP",
              "p14MH.1rIN_B14", "p14MH.1rIN_T41")

nameDF.2rIN = c("p14MH.2rIN_B0", "p14MH.2rIN_T0",
                "p14MH.2rIN_preREP", "p14MH.2rIN_ACTP",
                "p14MH.2rIN_B14", "p14MH.2rIN_T41")

```



```{r make statistic DF, echo=TRUE}
# Create DF for clone statistics
# DEV: need to create function to extract numbers from clonotypeCount & then add them to new df cloneStatDF

cloneStatDF <- data.frame(sample = nameDF,
                          all = c(430520, 7499, 1822, 10130, 12794, 28306),
                          IN = c(422319, 6956, 1690, 9452, 11829, 26746),
                          OUT = c(704, 246, 95, 318, 410, 523),
                          STOP = c(7439, 279, 35, 324, 531, 993),
                          AMB = c(58, 18, 2, 36, 24, 44),
                          uniq_nucSeq = c(281569, 5485, 1641, 8264, 8960, 19190),
                          uniq_peptide = c(267115, 5117, 1539, 7792, 8294, 18140)
                         )

cloneStatDF <- cloneStatDF %>%
  pivot_longer(!sample, names_to = "type", values_to = "count") %>% 
  mutate(log2_count = log2(count))

cloneStatDF

```

```{r make statistic DF.1r, echo=TRUE}
# Create DF.1r for clone statistics
# DEV: need to create function to extract numbers from clonotypeCount & then add them to new df cloneStatDF.1r

cloneStatDF.1r <- data.frame(sample = nameDF.1r,
                          all = c(230302, 5465, 1310, 5086, 6989, 17905),
                          IN = c(223619, 5153, 1224, 4692, 6470, 16975),
                          OUT = c(16, 51, 52, 107, 70, 66),
                          STOP = c(6609, 243, 32, 254, 428, 820),
                          AMB = c(58, 18, 2, 33, 21, 44),
                          uniq_nucSeq = c(186813, 4626, 1248, 4680, 6117, 14966),
                          uniq_peptide = c(180395, 4476, 1190, 4516, 5897, 14560)
                         )

cloneStatDF.1r <- cloneStatDF.1r %>%
  pivot_longer(!sample, names_to = "type", values_to = "count") %>% 
  mutate(log2_count = log2(count))

cloneStatDF.1r

```

```{r make statistic DF.1rIN, echo=TRUE}
# Create DF.1rIN for clone statistics
# DEV: need to create function to extract numbers from clonotypeCount & then add them to new df cloneStatDF.1rIN

cloneStatDF.1rIN <- data.frame(sample = nameDF.1rIN,
                               all.1rIN = c(223619, 5153, 1224, 4692, 6470, 16975),
                               uniq_nucSeq.1rIN = c(180209, 4317, 1162, 4291, 5609, 14057),
                               uniq_peptide.1rIN = c(173814, 4217, 1155, 4236, 5462, 13716)
                              )

cloneStatDF.1rIN <- cloneStatDF.1rIN %>%
  pivot_longer(!sample, names_to = "type", values_to = "count") %>% 
  mutate(log2_count = log2(count))

cloneStatDF.1rIN

```


## Slide with Plot

```{r plot clone types of all clones}

#source("barplot_cloneStat.R")
#statDfName = c("cloneStatDF", "cloneStatDF.1r", "cloneStatDF.1rIN")

pdf("./14MH/log2Count_of_cloneStatDF.pdf", width = 14, height = 4)
  ggplot(cloneStatDF, aes(fill = type, x = sample, y = log2_count)) + 
    geom_bar(position="dodge", stat="identity") +
    geom_text(aes(label = count), 
              hjust = -0.02, vjust = -0.7, 
              angle = 45, size = 3,
              position = position_dodge(0.9)) +
    ylim(0, max(cloneStatDF$log2_count)*1.2) +
    ggtitle("Counts of all clonotypes") +
    theme(axis.text=element_text(size=12, face = "bold")) +
    theme(legend.position = "bottom")
dev.off()

```


```{r plot clone types of 1r clones}

#source("barplot_cloneStat.R")
#statDfName = c("cloneStatDF", "cloneStatDF.1r", "cloneStatDF.1rIN")

cloneStatDF.1r$sample <- factor(cloneStatDF.1r$sample, 
                                levels = c("p14MH.1r_B0", "p14MH.1r_T0",
                                           "p14MH.1r_preREP", "p14MH.1r_ACTP",
                                           "p14MH.1r_B14", "p14MH.1r_T41"))

pdf("./14MH/log2Count_of_cloneStatDF.1r.pdf", width = 14, height = 4)
  ggplot(cloneStatDF.1r, aes(fill = type, x = sample, y = log2_count)) + 
    geom_bar(position="dodge", stat="identity") +
    geom_text(aes(label = count), 
              hjust = -0.02, vjust = -0.7, 
              angle = 45, size = 3,
              position = position_dodge(0.9)) +
    ylim(0, max(cloneStatDF.1r$log2_count)*1.2) +
    ggtitle("Counts of 1-read clonotypes") +
    theme(axis.text=element_text(size=12, face = "bold")) +
    theme(legend.position = "bottom")
dev.off()

```


```{r plot clone types of 1rIN clones}

#source("barplot_cloneStat.R")
#statDfName = c("cloneStatDF", "cloneStatDF.1r", "cloneStatDF.1rIN")

cloneStatDF.1rIN$sample <- factor(cloneStatDF.1rIN$sample, 
                                levels = c("p14MH.1rIN_B0", "p14MH.1rIN_T0",
                                           "p14MH.1rIN_preREP", "p14MH.1rIN_ACTP",
                                           "p14MH.1rIN_B14", "p14MH.1rIN_T41"))

pdf("./14MH/Count_of_cloneStatDF.1rIN.pdf", width = 10, height = 6)
  ggplot(cloneStatDF.1rIN, aes(fill = type, x = sample, y = count)) + 
    geom_bar(position="dodge", stat="identity") +
    geom_text(aes(label = count), 
              hjust = -0.02, vjust = -0.7, 
              angle = 45, size = 3,
              position = position_dodge(0.9)) +
    ylim(0, max(cloneStatDF.1rIN$count)*1.2) +
    ggtitle("Counts of 1-read IN-frame clonotypes") +
    theme(axis.text=element_text(size=8, face = "bold")) +
    theme(legend.position = "right")
dev.off()

```

# DEV: fix the following for loop
for (i in as.data.frame(statDfName)) {
  pdf(paste0("log2Count_of_", i, ".pdf"), width = 11, height = 4)
  barStat(i)
}


```{r, separate sub DFs into global environment, echo=FALSE}

list2env(setNames(list.DF, nameDF), .GlobalEnv)
#list2env(setNames(list.DF.1r, nameDF.1r), .GlobalEnv)
list2env(setNames(list.DF.IN, nameDF.IN), .GlobalEnv)
list2env(setNames(list.DF.1rIN, nameDF.1rIN), .GlobalEnv)
#list2env(setNames(list.DF.2rIN, nameDF.2rIN), .GlobalEnv)

```




## DEV!!
```{r, tracing 1read clones}
source("extract_1rIN.B0_persisted_in_all_INframe.R")
trace_1rIN.B0 <- lapply(list.DF.IN, extract_IN_with_1r.B0)

source("extract_1rIN.T0_persisted_in_all_INframe.R")
trace_1rIN.T0 <- lapply(list.DF.IN, extract_IN_with_1r.T0)

```


```{r, join all trace_B0 dataframes }

nameTR <- c("trace_B0.T0", "trace_B0.preREP", "trace_B0.ACTP", "trace_B0.B14", "trace_B0.T41")

source("join_trace.R")

  trace_B0.T0b <- left_join(trace_B0.T0, p14MH.1_B0,
                       by = "CDR3_sequence",
                       suffix = c(".T0", ".B0"))

# the following for loop does NOT work! 
for (i in nameTR) {
  join_trace(i)
}

```



```{r subset dataset for clones with only 1 read, echo=FALSE}

# add "clone" column to each dataframe of the 1read list 
# and separate the dataframes of list.DF.1read & list.DF.IN

add_cloneID <- function(spl,...){
  spl <- spl %>%
    unite("clone", c(TRBV, CDR3_aaseq, TRBJ), remove = F)
}

list.DF.1read <- lapply(list.DF.1read, add_cloneID)
```




